---
title: "Group 4 Mpox model"
author: "Group 4 members"
date: "2024-09-26"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Load the necessary libraries

```{r}

library(deSolve)
library(ggplot2)
library(reshape2)
library(MASS)

```






# Loading the dataset

```{r}
cases<-read.csv("C:/Users/bbrsh/Desktop/Shakira/Shakira/Infectious diseases modelling opportunity Kwame Nkuruma/Hands On/Group 4 things/MPox_docs/Shakira/mpox_2024.csv")
observed_data<-cases$new_cases #Extract the weekly case counts

```



```{r}
dir.create("plots", showWarnings = FALSE)

```




```{r}

# Define a function to save plots
save_plot <- function(plot, filename) {
  ggsave(filename, plot = plot, path = "C:/Users/bbrsh/Desktop/Shakira/Shakira/Infectious diseases modelling opportunity Kwame Nkuruma/Hands On/Group_4/Plots", width = 8, height = 6)
}

# Define base parameters
base_params <- c(
  beta_s = (0.25 * 2),  
  beta_ns = 0.25,         
  gamma_I = 0.6,              
  sigma = 0.3,                
  gamma_Id = 0.8,             
  rho_I = 0.065,                
  rho_Id = 0.03               
)

# Define initial state
initial_state <- c(
  S = 105e3,    # Susceptible
  I_s = 5,     # Infectious via the sexual route
  I_ns = 20,    # Infectious via other routes
  Id = 8,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 57 weeks
times <- seq(1, 57, by = 1)

# Define the model function
mpox_model <- function(time, state, parms) {
  with(as.list(c(state, parms)), {
    N <- S + I_s + I_ns + Id + R + D
    
    dS <- -beta_s * S * I_s / N - beta_ns * S * I_ns / N
    dI_s <- beta_s * S * I_s / N - gamma_I * I_s - sigma * I_s - rho_I * I_s
    dI_ns <- beta_ns * S * I_ns / N - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
    dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
    dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
    dD <- rho_I * (I_s + I_ns) + rho_Id * Id
    
    return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
  })
}

# Run the simulation
simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = base_params)
output <- as.data.frame(simulation)
colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")

# Plot the results
plot1 <- ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious_Sexual, color = "Infectious_Sexual"), size = 1) +
  geom_line(aes(y = Infectious_NonSexual, color = "Infectious_NonSexual"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  geom_line(aes(y = Dead, color = "Dead"), size = 1) +
  labs(title = "SIRD Model Simulation", x = "Time (weeks)", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead"),
                     values = c("blue", "red", "purple", "orange", "green", "black")) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Save the first plot
save_plot(plot1, "sird_model_simulation.png")


# Define cost function for parameter estimation
cost_function <- function(pars, observed_data) {
  params <- base_params
  params["beta_ns"] <- pars[1]
  params["sigma"] <- pars[2]

  # Simulate the model
  simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params)
  
  # Extract the number of infected individuals (Id)
  simulated_Id <- simulation[, "Id"]
  
  # Calculate the sum of squared errors
  return(sum((observed_data - simulated_Id)^2))
}

# Parameter estimation setup
lower_bounds <- c(beta_ns = 0.1, sigma = 0.1)
upper_bounds <- c(beta_ns = 0.5, sigma = 1)
initial_guesses <- c(beta_ns = 0.25, sigma = 0.3)

# Perform parameter estimation using optim with bounds
fit <- optim(par = initial_guesses, fn = cost_function, observed_data = observed_data, 
             method = "L-BFGS-B", lower = lower_bounds, upper = upper_bounds, hessian = TRUE)

# Extract estimated parameters
estimated <- fit$par
names(estimated) <- c("beta_ns", "sigma")

# Update the base parameters with estimated values
params <- base_params
params["beta_ns"] <- estimated["beta_ns"]
params["beta_s"] <- 2 * estimated["beta_ns"]
params["sigma"] <- estimated["sigma"]

# Run the simulation again with estimated parameters
simulation_updated <- ode(y = initial_state, times = times, func = mpox_model, parms = params)
predicted_Id <- simulation_updated[, "Id"]
comparison_df <- data.frame(
  time = times,
  Observed = observed_data,
  Predicted = predicted_Id
)

# Plot observed vs predicted
plot2 <- ggplot(data = comparison_df, aes(x = time)) +
  geom_line(aes(y = Observed, color = "Observed"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1, linetype = "dashed") +
  labs(title = "Observed vs. Predicted Detected Cases", x = "Time (weeks)", y = "Detected Cases") +
  scale_color_manual("", values = c("Observed" = "red", "Predicted" = "blue")) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Save the second plot
save_plot(plot2, "observed_vs_predicted.png")

# Simulate projection
project_time <- seq(0, max(times) + 52)  # 1 year more
projection <- ode(y = initial_state, times = project_time, func = mpox_model, parms = params)
projection_df <- as.data.frame(projection)

# Create a date vector for the projection period
start_date <- as.Date("2023-08-20")
projection_date_vector <- seq.Date(from = start_date, by = "week", length.out = length(project_time))

# Plot the projection with dates on the x-axis
plot3 <- ggplot(data = projection_df, aes(x = projection_date_vector, y = Id)) +
  geom_line(color = "red") +
  labs(title = "Projected Detected Infected Humans (Id)", x = "Date", y = "Detected Cases") +
  theme_minimal()

# Overlay observed data on the projection graph
observed_df <- data.frame(Date = projection_date_vector[1:length(observed_data)], Observed = observed_data)
plot3 <- plot3 + geom_point(data = observed_df, aes(x = Date, y = Observed), color = "blue") +
  scale_x_date(date_labels = "%b %Y") +
  theme(legend.position = "topright")

# Save the projection plot
save_plot(plot3, "projected_cases.png")



```



##############################################################################################

# Let's add an intervention e.g contact tracing, quarantine etc

# First run the model with the estimates estimates

```{r}

# Define base parameters
base_params <- c(
  beta_s = (0.5 * 2),  # Transmission rate via the sexual route
  beta_ns = 0.5,       # Transmission rate via other non-sexual routes
  gamma_I = 0.6,       # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.1910888,   # Detection rate
  gamma_Id = 0.8,      # Recovery rate for those who were detected
  rho_I = 0.065,       # Mortality rate for the infected people in the community
  rho_Id = 0.03        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e3,    # Susceptible
  I_s = 5,      # Infectious via the sexual route
  I_ns = 20,    # Infectious via other routes
  Id = 8,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 57 weeks
times <- seq(1, 57, by = 1)

# Define the model function
mpox_model <- function(time, state, parms) {
  with(as.list(c(state, parms)), {
    N <- S + I_s + I_ns + Id + R + D
    
    dS <- -beta_s * S * I_s / N - beta_ns * S * I_ns / N
    dI_s <- beta_s * S * I_s / N - gamma_I * I_s - sigma * I_s - rho_I * I_s
    dI_ns <- beta_ns * S * I_ns / N - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
    dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
    dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
    dD <- rho_I * (I_s + I_ns) + rho_Id * Id
    
    return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
  })
}

# Run the simulation
simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = base_params)

# Convert the output to a data frame for analysis and plotting
output <- as.data.frame(simulation)

# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")

# Calculate total infected
output$Total_Infected <- output$Infectious_Sexual + output$Infectious_NonSexual

# Display the output
print(output)

# Plot only the total infected
plot2 = ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Total_Infected, color = "Total Infected"), size = 1) +  # Plot Total Infected
  labs(title = "Total Infected Population Over Time", x = "Time (weeks)", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Total Infected"),
                     values = c("brown")) +  # Set color for Total Infected
  theme_minimal() +
  theme(legend.position = "bottom")




# Save the first plot
save_plot(plot2, "sird_model_simulation_with_estimated_parameters.png")


```





# Now, let's introduce an intervention like contact tracing. Contact tracing reduces the transmission rate by quickly removing infected individuals. Using 
(1âˆ’mð‘¢)*beta_s and (1âˆ’mu)*beta_ns , we'll examine mu values of 0, 0.01, 0.04, and 0.08, and plot the total number of infected individuals for each mu.

```{r}

# Define base parameters
base_params <- c(
  beta_s = (0.5 * 2),  # Transmission rate via the sexual route
  beta_ns = 0.5,       # Transmission rate via other non-sexual routes
  gamma_I = 0.6,       # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.1910888,   # Detection rate
  gamma_Id = 0.8,      # Recovery rate for those who were detected
  rho_I = 0.065,       # Mortality rate for the infected people in the community
  rho_Id = 0.03        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e3,    # Susceptible
  I_s = 5,      # Infectious via the sexual route
  I_ns = 20,    # Infectious via other routes
  Id = 8,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 57 weeks
times <- seq(1, 57, by = 1)

# Define the model function with contact tracing effect
mpox_model <- function(time, state, parms) {
  with(as.list(c(state, parms)), {
    N <- S + I_s + I_ns + Id + R + D
    
    # Calculate the modified transmission rates
    mu <- parms["mu"]  # Intervention effectiveness
    effective_beta_s <- (1 - mu) * beta_s
    effective_beta_ns <- (1 - mu) * beta_ns
    
    dS <- -effective_beta_s * S * I_s / N - effective_beta_ns * S * I_ns / N
    dI_s <- effective_beta_s * S * I_s / N - gamma_I * I_s - sigma * I_s - rho_I * I_s
    dI_ns <- effective_beta_ns * S * I_ns / N - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
    dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
    dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
    dD <- rho_I * (I_s + I_ns) + rho_Id * Id
    
    return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
  })
}

# Values for mu to test
mu_values <- c(0,0.01, 0.04, 0.08)

# Prepare an empty list to store simulation results
simulation_results <- list()

# Run simulations for different values of mu
for (mu in mu_values) {
  # Add mu to parameters
  params_with_mu <- c(base_params, mu = mu)
  
  # Run the simulation
  simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params_with_mu)
  
  # Convert the output to a data frame for analysis and plotting
  output <- as.data.frame(simulation)
  
  # Rename columns for clarity
  colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")
  
  # Calculate total infected
  output$Total_Infected <- output$Infectious_Sexual + output$Infectious_NonSexual
  
  # Store results with mu as a list entry
  simulation_results[[paste0("mu_", mu)]] <- output
}

# Combine results into one data frame for plotting
combined_results <- do.call(rbind, lapply(names(simulation_results), function(name) {
  data <- simulation_results[[name]]
  data$mu <- name  # Add the mu value as a column
  return(data)
}))


# Plot total infected for each mu value
plot3 = ggplot(data = combined_results, aes(x = time, y = Total_Infected, color = mu)) +
  geom_line(size = 1) +  # Plot Total Infected for each mu
  labs(title = "Total Infected Population Over Time with Contact Tracing", 
       x = "Time (weeks)", 
       y = "Population") +
  scale_color_manual("",
                     breaks = c("mu_0", "mu_0.01", "mu_0.04", "mu_0.08"),  # Correct breaks for mu
                     labels = c("Contact Tracing (Î¼ = 0)", "Contact Tracing (Î¼ = 0.01)", 
                                "Contact Tracing (Î¼ = 0.04)", "Contact Tracing (Î¼ = 0.08)"),  # Update labels
                     values = c("brown", "orange", "green", "purple")) +  # Set colors for mu values
  theme_minimal() +
  theme(legend.position = "bottom")

# Save the first plot
save_plot(plot3, "Total_number_infected_per_different_contact_tracing_rates.png")



```



# Let's also examine an intervention like quarantine. This also reduces the beta values. We'll consider quarantine rates of 0, 0.01, 0.04, and 0.08.

```{r}

# Define base parameters
base_params <- c(
  beta_s = (0.5 * 2),  # Transmission rate via the sexual route
  beta_ns = 0.5,       # Transmission rate via other non-sexual routes
  gamma_I = 0.6,       # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.1910888,   # Detection rate
  gamma_Id = 0.8,      # Recovery rate for those who were detected
  rho_I = 0.065,       # Mortality rate for the infected people in the community
  rho_Id = 0.03        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e3,    # Susceptible
  I_s = 5,      # Infectious via the sexual route
  I_ns = 20,    # Infectious via other routes
  Id = 8,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 57 weeks
times <- seq(1, 57, by = 1)

# Define the model function with quarantine effect
mpox_model <- function(time, state, parms) {
  with(as.list(c(state, parms)), {
    N <- S + I_s + I_ns + Id + R + D
    
    # Calculate the modified transmission rates
    mu <- parms["mu"]  # Intervention effectiveness
    effective_beta_s <- (1 - mu) * beta_s
    effective_beta_ns <- (1 - mu) * beta_ns
    
    dS <- -effective_beta_s * S * I_s / N - effective_beta_ns * S * I_ns / N
    dI_s <- effective_beta_s * S * I_s / N - gamma_I * I_s - sigma * I_s - rho_I * I_s
    dI_ns <- effective_beta_ns * S * I_ns / N - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
    dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
    dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
    dD <- rho_I * (I_s + I_ns) + rho_Id * Id
    
    return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
  })
}

# Values for mu to test
mu_values <- c(0,0.01, 0.04, 0.08)

# Prepare an empty list to store simulation results
simulation_results <- list()

# Run simulations for different values of mu
for (mu in mu_values) {
  # Add mu to parameters
  params_with_mu <- c(base_params, mu = mu)
  
  # Run the simulation
  simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params_with_mu)
  
  # Convert the output to a data frame for analysis and plotting
  output <- as.data.frame(simulation)
  
  # Rename columns for clarity
  colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")
  
  # Calculate total infected
  output$Total_Infected <- output$Infectious_Sexual + output$Infectious_NonSexual
  
  # Store results with mu as a list entry
  simulation_results[[paste0("mu_", mu)]] <- output
}

# Combine results into one data frame for plotting
combined_results <- do.call(rbind, lapply(names(simulation_results), function(name) {
  data <- simulation_results[[name]]
  data$mu <- name  # Add the mu value as a column
  return(data)
}))


# Plot total infected for each mu value
plot4 = ggplot(data = combined_results, aes(x = time, y = Total_Infected, color = mu)) +
  geom_line(size = 1) +  # Plot Total Infected for each mu
  labs(title = "Total Infected Population Over Time with quarantine", 
       x = "Time (weeks)", 
       y = "Population") +
  scale_color_manual("",
                     breaks = c("mu_0", "mu_0.01", "mu_0.04", "mu_0.08"),  # Correct breaks for mu
                     labels = c("quarantine (Î¼ = 0)", "quarantine (Î¼ = 0.01)", 
                                "quarantine (Î¼ = 0.04)", "quarantine (Î¼ = 0.08)"),  # Update labels
                     values = c("brown", "orange", "green", "purple")) +  # Set colors for mu values
  theme_minimal() +
  theme(legend.position = "bottom")


# Save the first plot
save_plot(plot4, "Total_number_infected_per_different_quarantine_rates.png")



```



# Let's now compare contact tracing with a rate of 0.02 on its own and contact tracing with a rate of 0.04 on its own, to the combination of both contact tracing (0.02) and quarantine (0.04). Both impact the beta values.

```{r}
# Define base parameters
base_params <- c(
  beta_s = (0.5 * 2),  # Transmission rate via the sexual route
  beta_ns = 0.5,       # Transmission rate via other non-sexual routes
  gamma_I = 0.6,       # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.1910888,   # Detection rate
  gamma_Id = 0.8,      # Recovery rate for those who were detected
  rho_I = 0.065,       # Mortality rate for the infected people in the community
  rho_Id = 0.03        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e3,    # Susceptible
  I_s = 5,      # Infectious via the sexual route
  I_ns = 20,    # Infectious via other routes
  Id = 8,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 57 weeks
times <- seq(1, 57, by = 1)

# Define the model function with both contact tracing and quarantine effects
mpox_model_combined <- function(time, state, parms) {
  with(as.list(c(state, parms)), {
    N <- S + I_s + I_ns + Id + R + D
    
    # Calculate the modified transmission rates
    contact_tracing <- parms["contact_tracing"]
    quarantine <- parms["quarantine"]
    
    effective_beta_s <- (1 - contact_tracing) * (1 - quarantine) * beta_s
    effective_beta_ns <- (1 - contact_tracing) * (1 - quarantine) * beta_ns
    
    dS <- -effective_beta_s * S * I_s / N - effective_beta_ns * S * I_ns / N
    dI_s <- effective_beta_s * S * I_s / N - gamma_I * I_s - sigma * I_s - rho_I * I_s
    dI_ns <- effective_beta_ns * S * I_ns / N - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
    dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
    dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
    dD <- rho_I * (I_s + I_ns) + rho_Id * Id
    
    return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
  })
}

# Values for interventions to test
intervention_values <- c(0.02,0.02, 0.04)  # Contact tracing and quarantine rates of 0.02

# Prepare an empty list to store simulation results
simulation_results <- list()

# Run simulations for different scenarios: no intervention, contact tracing only, quarantine only, and both combined
scenarios <- list(
  "No_Intervention" = c(contact_tracing = 0, quarantine = 0),
  "Contact_Tracing_0.02" = c(contact_tracing = 0.02, quarantine = 0),
  "Quarantine_0.04" = c(contact_tracing = 0, quarantine = 0.04),
  "Both_Combined_0.02_and_0.04" = c(contact_tracing = 0.02, quarantine = 0.04)
)

for (scenario in names(scenarios)) {
  # Add the intervention values to parameters
  params_with_intervention <- c(base_params, scenarios[[scenario]])
  
  # Run the simulation
  simulation <- ode(y = initial_state, times = times, func = mpox_model_combined, parms = params_with_intervention)
  
  # Convert the output to a data frame for analysis and plotting
  output <- as.data.frame(simulation)
  
  # Rename columns for clarity
  colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")
  
  # Calculate total infected
  output$Total_Infected <- output$Infectious_Sexual + output$Infectious_NonSexual
  
  # Store results with the scenario name as a list entry
  simulation_results[[scenario]] <- output
}

# Combine results into one data frame for plotting
combined_results <- do.call(rbind, lapply(names(simulation_results), function(name) {
  data <- simulation_results[[name]]
  data$scenario <- name  # Add the scenario name as a column
  return(data)
}))

# Plot total infected for each scenario
plot5 = ggplot(data = combined_results, aes(x = time, y = Total_Infected, color = scenario)) +
  geom_line(size = 1) +  # Plot Total Infected for each scenario
  labs(title = "Total Infected Population Over Time with Various Interventions", 
       x = "Time (weeks)", 
       y = "Population") +
  scale_color_manual("",
                     breaks = c("No_Intervention", "Contact_Tracing_0.02", "Quarantine_0.04", "Both_Combined_0.02_and_0.04"),  
                     labels = c("No Intervention (Î¼ = 0)", 
                                "Contact Tracing (Î¼ = 0.02)", 
                                "Quarantine (Î¼ = 0.04)", 
                                "Both Combined (Î¼ = 0.02 and Î¼ = 0.04)"),  
                     values = c("black", "blue", "red", "purple")) +  # Set colors for scenarios
  theme_minimal() +
  theme(legend.position = "bottom")


# Save the first plot
save_plot(plot5, "Total_number_infected_per_different_interventions.png")





```


















