---
title: "Group 4 Mathematical model for Mpox"
author: "Group 4 members"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load the necessary libraries

```{r}

library(deSolve)
library(ggplot2)
library(reshape2)
library(MASS)

```



# Loading the dataset

```{r}
cases<-read.csv("C:/Users/bbrsh/Desktop/Shakira/Shakira/Infectious diseases modelling opportunity Kwame Nkuruma/Hands On/Group_4/MPox_docs/Mpox_data/mpox_2024.csv")
observed_data<-cases$new_cases #Extract the weekly case counts

```


# Simple SIR Model


```{r}
# Define base parameters 
base_params <- c(
  beta = 0.05,      # Transmission rate 
  gamma = 0.7       # Recovery rate for I 
  
)

# Define initial state
initial_state <- c(
  S = 105e6,        # Susceptible
  I = 1,          # Infectious
  R = 0         # Recovered
)

# Define the time sequence for 57 weeks
times <- seq(0, 57, by = 1)

# Function to run the model
run_model <- function(params) {
  model <- function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta * S * I
      dI <- beta * S * I - gamma * I 
      dR <- gamma * I 
      
      return(list(c(dS, dI, dR)))
    })
  }
  
  # Run the model
  output <- ode(y = initial_state, times = times, func = model, parms = params)
  
  return(as.data.frame(output))
}

# Run the model
output <- run_model(base_params)

output


# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious", "Recovered")

# Plot the results
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious, color = "Infectious"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  labs(title = "SIR Model Simulation", x = "Days", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious", "Recovered"),
                     values = c("blue", "red", "green")) +
  theme_minimal() +
  theme(legend.position = "bottom")


```






```{r}
# Define base parameters 
base_params <- c(
  beta = 0.05,      # Transmission rate 
  gamma = 0.07       # Recovery rate for I 
  
)

# Define initial state
initial_state <- c(
  S = 105e6,        # Susceptible
  I = 20,          # Infectious
  R = 0         # Recovered
)

# Define the time sequence for the 57 weeks
times <- seq(0, 57, by = 1)

# Function to run the model
run_model <- function(params) {
  model <- function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta * S * I
      dI <- beta * S * I - gamma * I 
      dR <- gamma * I 
      
      return(list(c(dS, dI, dR)))
    })
  }
  
  # Run the model
  output <- ode(y = initial_state, times = times, func = model, parms = params)
  
  return(as.data.frame(output))
}

# Run the model
output <- run_model(base_params)

output


# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious", "Recovered")

# Plot the results
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious, color = "Infectious"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  labs(title = "SIR Model Simulation", x = "Days", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious", "Recovered"),
                     values = c("blue", "red", "green")) +
  theme_minimal() +
  theme(legend.position = "bottom")


```


############################################################################################################
# Let's consider the fact that not everyone is detected



```{r}

# Define base parameters
base_params <- c(
  beta = 0.05,    # Transmission rate
  gamma_I = 0.7,    # Recovery rate for general infectious population
  sigma = 0.05,       # Detection rate
  gamma_Id = 0.14   # Recovery rate for detected population
)

# Define initial state
initial_state <- c(
  S = 105e6,    # Susceptible
  I = 20,     # Infectious
  Id = 2,     # Detected population
  R = 0       # Recovered
)

# Define the time sequence (57 weeks)
times <- seq(0, 57, by = 1)  # Replace with appropriate time frame

# Define the ODE system
mpox_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # ODEs based on the flowchart parameters
    dS <- -beta * S * I
    dI <- beta * S * I - gamma_I * I - sigma * I
    dId <- sigma * I - gamma_Id * Id
    dR <- gamma_I * I + gamma_Id * Id
    
    return(list(c(dS, dI, dId, dR)))
  })
}

# Run the model
output <- ode(y = initial_state, times = times, func = mpox_model, parms = base_params)

# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious", "Detected", "Recovered")

# Check the output to ensure it's being generated (Optional)
print(head(output))

# Plot the results
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious, color = "Infectious"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  labs(title = "SIR Model Simulation", x = "Days", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious", "Detected", "Recovered"),
                     values = c("blue", "red", "brown", "green")) +
  theme_minimal() +
  theme(legend.position = "bottom")

############################################################################################################
# Parameter estimation for sigma using least squared method

# Define cost function for parameter estimation
cost_function <- function(pars, observed_data) {
  # Create a copy of base_params and update sigma
  params <- base_params
  params["sigma"] <- pars[1]
  
  # Simulate the model
  simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params)
  
  # Extract the number of detected individuals (Id)
  simulated_Id <- simulation[, "Id"]
  
  # Calculate the sum of squared errors
  return(sum((observed_data - simulated_Id)^2))
}
  
# Define lower and upper bounds for the parameters
lower_bounds <- c(sigma = 0.01)
upper_bounds <- c(sigma = 1)

# Initial guesses for the parameters to be estimated
initial_guesses <- lower_bounds

# Perform parameter estimation using optim with bounds
fit <- optim(par = initial_guesses, fn = cost_function, observed_data = observed_data, 
             method = "L-BFGS-B", lower = lower_bounds, upper = upper_bounds, hessian = TRUE)

# Extract the estimated parameters
estimated <- fit$par
##names(estimated) <- c("sigma")
print(estimated)

# Compute confidence intervals
hessian_matrix <- fit$hessian
cov_matrix <- solve(hessian_matrix)  # Invert the Hessian matrix to get the covariance matrix
std_errors <- sqrt(diag(cov_matrix))  # Standard errors of the parameters
conf_intervals <- cbind(estimated - 1.96 * std_errors, estimated + 1.96 * std_errors)
print(conf_intervals)


# Recreate the base parameters with the estimated sigma
params <- base_params
params["sigma"] <- estimated["sigma"]

#params["sigma"] <- estimated["sigma"]




```


# Parameter estimation using Maximum likelihood estimation (MLE)

```{r}

# Define log-likelihood function for MLE
log_likelihood <- function(pars, observed_data) {
  params <- base_params
  params["sigma"] <- pars[1]
  
  simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params)
  simulated_Id <- simulation[, "Id"]
  
  # Assuming Gaussian noise, the log-likelihood is proportional to the negative SSE
  neg_log_likelihood <- -sum(dnorm(observed_data, mean = simulated_Id, sd = sd(observed_data - simulated_Id), log = TRUE))
  return(neg_log_likelihood)
}

# Perform MLE using optim
mle_fit <- optim(par = initial_guesses, fn = log_likelihood, observed_data = observed_data, 
                 method = "L-BFGS-B", lower = lower_bounds, upper = upper_bounds, hessian = TRUE)

# Extract and print the estimated sigma
mle_estimated <- mle_fit$par
print(mle_estimated)


```





# Rerunning the model with the estimated sigma

```{r}

# Define base parameters
base_params <- c(
  beta = 0.05,    # Transmission rate
  gamma_I = 0.7,    # Recovery rate for general infectious population
  sigma = 0.01,       # Detection rate
  gamma_Id = 0.14   # Recovery rate for detected population
)

# Define initial state
initial_state <- c(
  S = 105e6,    # Susceptible
  I = 20,     # Infectious
  Id = 2,     # Detected population
  R = 0       # Recovered
)

# Define the time sequence (57 weeks)
times <- seq(0, 57, by = 1)  # Replace with appropriate time frame

# Define the ODE system
mpox_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # ODEs based on the flowchart parameters
    dS <- -beta * S * I
    dI <- beta * S * I - gamma_I * I - sigma * I
    dId <- sigma * I - gamma_Id * Id
    dR <- gamma_I * I + gamma_Id * Id
    
    return(list(c(dS, dI, dId, dR)))
  })
}

# Run the model
output <- ode(y = initial_state, times = times, func = mpox_model, parms = base_params)

# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious", "Detected", "Recovered")

# Check the output to ensure it's being generated (Optional)
print(head(output))

# Plot the results
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious, color = "Infectious"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  labs(title = "SIR Model Simulation", x = "Days", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious", "Detected", "Recovered"),
                     values = c("blue", "red", "brown", "green")) +
  theme_minimal() +
  theme(legend.position = "bottom")


```


##########################################################################################################

```{r}
# Simulate again with new parameters
simulation <- ode(y = initial_state, times = times, func = mpox_model, parms = params)
simulation_df <- as.data.frame(simulation)

# Define start date for the timeline
start_date <- as.Date("2023-08-20")  # Example start date

# Create a date vector for the time points (based on weeks, as the time vector is weekly)
date_vector <- seq.Date(from = start_date, by = "week", length.out = length(times))
project_time <- seq(max(times),max(times)+52*1) # 1 year more (52 Weeks)
projection <- ode(y = initial_state, times = project_time, func = mpox_model, parms = params)
projection_df <- as.data.frame(projection)

# Create a date vector for the projection period
projection_date_vector <- seq.Date(from = max(date_vector), by = "week", length.out = length(project_time))

# Plot the observed and simulated results with dates on the x-axis
print(
  plot(date_vector, simulation_df$Ihd, type = "l", col = "red", ylab = "Detected Infected Humans (Id)", xlab = "Date", xaxt = "n")
)
# Overlay observed data (make sure 'observed_data' has the correct length)
# Ensure the lengths of observed data and date vector match
if (length(observed_data) > length(date_vector)) {
  observed_data <- observed_data[1:length(date_vector)]
} else if (length(observed_data) < length(date_vector)) {
  date_vector <- date_vector[1:length(observed_data)]
}

# Plot the simulated detected cases with dates on the x-axis
plot(date_vector, simulation_df$Detected, type = "l", col = "red", ylab = "Detected Infected Humans (Id)", xlab = "Date")

# Overlay observed data (ensure 'observed_data' and 'date_vector' have the same length)
points(date_vector, observed_data, col = "blue", pch = 16)

# Add a legend to differentiate between simulated and observed data
legend("topright", legend = c("Simulated Detected", "Observed Data"), col = c("red", "blue"), lty = 1, pch = c(NA, 16))



#points(date_vector, observed_data, col = "blue", pch = 16)

# Format the x-axis to show dates
axis.Date(1, at = date_vector, format = "%b %Y")  # Custom format: "Month Year"
legend("topleft", legend = c("Simulated Id", "Observed Data"), col = c("red", "blue"), lty = c(1, NA), pch=c(NA,16))

# Plot the projection with dates on the x-axis
print(
  plot(projection_date_vector, projection_df$Id, type = "l", col = "red", ylab = "Projected detected Infected Humans (Id)", xlab = "Date", xaxt = "n")
)
axis.Date(1, at = projection_date_vector, format = "%b %Y")




```



#########################################################################################################

```{r}

#################### Projection ( 2023 to 2025)
#################### 

project_time <- seq(0,max(times)+52*1) # 1 year more (52 Weeks)
projection <- ode(y = initial_state, times = project_time, func = mpox_model, parms = params)
projection_df <- as.data.frame(projection)

# Create a date vector for the projection period
projection_date_vector <- seq.Date(from = start_date, by = "week", length.out = length(project_time))


#Plot the projection with dates on the x-axis
print(
  plot(projection_date_vector, projection_df$Id, type = "l", col = "red", ylab = "Projected Detected Infected Humans (Id)", xlab = "Date", xaxt = "n")
)
# Overlay observed data on the projection graph (observed data only goes to the current time)
points(date_vector, observed_data, col = "blue", pch = 16)  # Observed data only goes up to current time
axis.Date(1, at = projection_date_vector, format = "%b %Y")
legend("topleft", legend = c("Projected MPOX Id", "Observed Data"), col = c("red", "blue"), lty = c(1, NA), pch=c(NA,16))

```






############################################################################################################
# Let's consider the fact that infected people die



```{r}

# Define base parameters 
base_params <- c(
  beta = 0.05,         # Transmission rate 
  gamma_I = 0.7,        # Recovery rate for those who were infectious and remained on the general population
  sigma = 0.01,            #Detection rate
  gamma_Id = 0.14,        # Recovery rate for thise who were detected
  rho_I = 0.049,           # Mortality rate for the infected people in the community
  rho_Id = 0.0455          # Mortality rate for the infected people that were detected and isolated
  
)

# Define initial state
initial_state <- c(
  S = 105e6,        # Susceptible
  I = 20,         # Infectious
  Id = 2,         #Detected population
  R = 0,         # Recovered
  D = 0         #Dead
)

# Define the time sequence for 37 weeks
times <- seq(0, 37, by = 1)

# Function to run the model
run_model <- function(params) {
  model <- function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta * S * I
      dI <- beta * S * I - gamma_I * I - sigma*I - rho_I*I
      dId <- sigma*I - gamma_Id *Id - rho_Id*Id
      dR <- gamma_I * I + gamma_Id *Id
      dD <- rho_I*I + rho_Id*Id
      
      return(list(c(dS, dI, dId, dR, dD)))
    })
  }
  
  # Run the model
  output <- ode(y = initial_state, times = times, func = model, parms = params)
  
  return(as.data.frame(output))
}

# Run the model
output <- run_model(base_params)

output


# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious","Detected", "Recovered","Dead")

# Plot the results
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious, color = "Infectious"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  geom_line(aes(y = Dead, color = "Dead"), size = 1) +
  labs(title = "SIR Model Simulation", x = "Days", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious", "Detected","Recovered","Dead"),
                     values = c("blue", "red","brown", "green","black")) +
  theme_minimal() +
  theme(legend.position = "bottom")

```



###########################################################################################################
# Adding the next transmission routes (sexual vs non-sexual)


```{r}


# Define base parameters
base_params <- c(
  beta_s = 0.05,         # Transmission rate via the sexual route
  beta_ns = 0.06,        # Transmission rate via other non-sexual routes
  gamma_I = 0.7,         # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.01,          # Detection rate
  gamma_Id = 0.14,       # Recovery rate for those who were detected
  rho_I = 0.049,         # Mortality rate for the infected people in the community
  rho_Id = 0.0455        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e6,    # Susceptible
  I_s = 9,      # Infectious via the sexual route
  I_ns = 11,    # Infectious via other routes
  Id = 2,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 37 weeks
times <- seq(0, 37, by = 1)

# Function to run the model
run_model <- function(params) {
  model <- function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta_s * S * I_s - beta_ns * S * I_ns
      dI_s <- beta_s * S * I_s - gamma_I * I_s - sigma * I_s - rho_I * I_s
      dI_ns <- beta_ns * S * I_ns - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
      dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
      dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
      dD <- rho_I * (I_s + I_ns) + rho_Id * Id
      
      return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
    })
  }
  
  # Run the model
  output <- ode(y = initial_state, times = times, func = model, parms = params)
  
  return(as.data.frame(output))
}

# Run the model
output <- run_model(base_params)

# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")

# Display the output
print(output)

# Plot the results
library(ggplot2)
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious_Sexual, color = "Infectious_Sexual"), size = 1) +
  geom_line(aes(y = Infectious_NonSexual, color = "Infectious_NonSexual"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  geom_line(aes(y = Dead, color = "Dead"), size = 1) +
  labs(title = "SEIRD Model Simulation", x = "Time (weeks)", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead"),
                     values = c("blue", "red", "purple", "orange", "green", "black")) +
  theme_minimal() +
  theme(legend.position = "bottom")



```



# Let us estimate the beta_s


```{r}

# Parameter estimation using least squared method

# Define cost function for parameter estimation
cost_function <- function(pars, observed_data) {
  # Create a copy of base_params and update the parameter to be estimated
  params <- base_params
  params["beta_s"] <- pars[1]  # Update beta_s with the estimated value
  
  # Simulate the model using updated parameters
  simulation <- ode(y = initial_state, times = times, func = function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta_s * S * I_s - beta_ns * S * I_ns
      dI_s <- beta_s * S * I_s - gamma_I * I_s - sigma * I_s - rho_I * I_s
      dI_ns <- beta_ns * S * I_ns - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
      dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
      dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
      dD <- rho_I * (I_s + I_ns) + rho_Id * Id
      
      return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
    })
  }, parms = params)
  
  simulation_df <- as.data.frame(simulation)
  
  # Extract the number of individuals who acquired the infection sexually (I_s)
  simulated_I_s <- simulation_df$I_s
  
  # Calculate the sum of squared errors (SSE)
  return(sum((observed_data - simulated_I_s)^2))
}

# Define lower and upper bounds for the parameters
lower_bounds <- c(beta_s = 0.01)
upper_bounds <- c(beta_s = 1)

# Initial guesses for the parameters to be estimated
initial_guesses <- c(beta_s = 0.05)

# Example observed data (this should be defined by you based on actual observations)
# observed_data <- c(...)  # Replace this with your real observed data

# Perform parameter estimation using optim with bounds and suppress warnings
fit <- suppressWarnings(
  optim(
    par = initial_guesses, 
    fn = cost_function, 
    observed_data = observed_data, 
    method = "L-BFGS-B", 
    lower = lower_bounds, 
    upper = upper_bounds, 
    hessian = TRUE
  )
)

# Display the fitted parameter estimates
print(fit$par)


# Compute confidence intervals
hessian_matrix <- fit$hessian
cov_matrix <- solve(hessian_matrix)  # Invert the Hessian matrix to get the covariance matrix
std_errors <- sqrt(diag(cov_matrix))  # Standard errors of the parameters
conf_intervals <- cbind(fit$par - 1.96 * std_errors, fit$par + 1.96 * std_errors)
print(conf_intervals)

params["beta_s"] <- fit$par["beta_s"]



```



```{r}

# Define base parameters
base_params <- c(
  beta_s = 0.01313275,         # Transmission rate via the sexual route
  beta_ns = 0.06,        # Transmission rate via other non-sexual routes
  gamma_I = 0.7,         # Recovery rate for those who were infectious and remained in the general population
  sigma = 0.01,          # Detection rate
  gamma_Id = 0.14,       # Recovery rate for those who were detected
  rho_I = 0.049,         # Mortality rate for the infected people in the community
  rho_Id = 0.0455        # Mortality rate for the infected people that were detected and isolated
)

# Define initial state
initial_state <- c(
  S = 105e6,    # Susceptible
  I_s = 9,      # Infectious via the sexual route
  I_ns = 11,    # Infectious via other routes
  Id = 2,       # Detected population
  R = 0,        # Recovered
  D = 0         # Dead
)

# Define the time sequence for 37 weeks
times <- seq(0, 37, by = 1)

# Function to run the model
run_model <- function(params) {
  model <- function(t, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta_s * S * I_s - beta_ns * S * I_ns
      dI_s <- beta_s * S * I_s - gamma_I * I_s - sigma * I_s - rho_I * I_s
      dI_ns <- beta_ns * S * I_ns - gamma_I * I_ns - sigma * I_ns - rho_I * I_ns
      dId <- sigma * (I_s + I_ns) - gamma_Id * Id - rho_Id * Id
      dR <- gamma_I * (I_s + I_ns) + gamma_Id * Id
      dD <- rho_I * (I_s + I_ns) + rho_Id * Id
      
      return(list(c(dS, dI_s, dI_ns, dId, dR, dD)))
    })
  }
  
  # Run the model
  output <- ode(y = initial_state, times = times, func = model, parms = params)
  
  return(as.data.frame(output))
}

# Run the model
output <- run_model(base_params)

# Rename columns for clarity
colnames(output) <- c("time", "Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead")

# Display the output
print(output)

# Plot the results
library(ggplot2)
ggplot(data = output, aes(x = time)) +
  geom_line(aes(y = Susceptible, color = "Susceptible"), size = 1) +
  geom_line(aes(y = Infectious_Sexual, color = "Infectious_Sexual"), size = 1) +
  geom_line(aes(y = Infectious_NonSexual, color = "Infectious_NonSexual"), size = 1) +
  geom_line(aes(y = Detected, color = "Detected"), size = 1) +
  geom_line(aes(y = Recovered, color = "Recovered"), size = 1) +
  geom_line(aes(y = Dead, color = "Dead"), size = 1) +
  labs(title = "SEIRD Model Simulation", x = "Time (weeks)", y = "Population") +
  scale_color_manual("", 
                     breaks = c("Susceptible", "Infectious_Sexual", "Infectious_NonSexual", "Detected", "Recovered", "Dead"),
                     values = c("blue", "red", "purple", "orange", "green", "black")) +
  theme_minimal() +
  theme(legend.position = "bottom")



```


















